The not so widening achievement gap: an interntational comparison of the evolution of the SES achievement gap from 2000-2015

### 2017-07-05 paper/paper.Rnw
Fixed shapes of two graphs

var_years %>%
  setNames(seq(2000, 2015, 3)) %>%
  enframer("name") %>%
  ggplot(aes(name, value)) +
  geom_col() +
  scale_y_continuous(name = "Standard deviation of raw test scores") +
  scale_x_discrete(name = NULL) +
  ggtitle("Standard deviation of test scores over time") +
  theme_few()
@
\caption{Standard deviation of test scores across all waves.}
\end{center}
\end{figure}

@@

complete_data_topbottom %>%
  bind_rows(complete_data_topmid) %>%
  bind_rows(complete_data_midbottom) %>%
  filter(country %in% c("United States", "Denmark", "France", "Germany"), type_test == "math") %>%
  select(wave, country, type_test, type, contains("math")) %>%
  mutate(type = factor(type,
                       levels = c("90th/10th SES gap", "90th/50th SES gap", "50th/10th SES gap"),
                       ordered = TRUE)) %>%
  gather(score, value, -(wave:type)) %>%
  separate(score, c("ses", "score"), sep = 2) %>%
  spread(score, value) %>%
  ggplot(aes(as.factor(wave), mean_math, group = ses, colour = ses)) +
  geom_errorbar(aes(ymin = lower_math, ymax = upper_math), width = 0.1) +
  geom_line(alpha = 0.4) +
  geom_point(size = 0.5, alpha = 0.4) +
  coord_cartesian(ylim = c(-1, 2)) +
  scale_y_continuous(name = "Standardized test scores (mean 0)") +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  scale_colour_discrete(name = NULL, labels = c("Low SES", "High SES")) +
  ggtitle("Evolution of the top/bottom groups by different gaps") +
  facet_grid(country ~ type) +
  theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        legend.position = "bottom")
@
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
<<avgperformance_inequality, include = T, out.height = '5in', out.width = '6.5in', fig.align = 'center'>>=
data_to_plot %>%
  ggplot(aes(scale(Mean), difference)) +
  geom_smooth(method = "lm") +
  geom_point() +
  scale_y_continuous(name = "90/10 achievement gap") +
  scale_x_continuous(name = "Average performance standardized") +
  ggtitle("Relationship between average performance and achievement gap") +
  facet_wrap(~ wave, nrow = 2) +
  theme_few()
@
\caption{Relationship between 90/10 achievement gap and average performance in mathematics}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
<<country_achievement_and_inequality, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
data_to_plot %>%
  mutate(Mean = scale(Mean),
         country = factor(country, levels = ordered_cnt, ordered = TRUE)) %>%
  gather(type, value, difference, Mean) %>%
  filter(country %in% countries) %>%
  ggplot(aes(as.character(wave), value, group = type, colour = type, shape = type)) +
  geom_point(alpha = 0.4) +
  geom_line(alpha = 0.4) +
  scale_y_continuous(name = "Standardized test scores (mean 0)") +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  scale_colour_discrete(name = NULL, labels = c("Achievement gap", "Avg performance")) +
  scale_shape_discrete(name = NULL, labels = c("Achievement gap", "Avg performance")) +
  ggtitle("Evolution of the top/bottom groups by different gaps") +
  facet_wrap(~ country, ncol = 5) +
  theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        legend.position = "bottom")
@
\caption{Time trends of achievement gap and average achiement between 2000 and 2015}
\end{center}
\end{figure}

>>>>

var_years %>%
  setNames(seq(2000, 2015, 3)) %>%
  enframer("name") %>%
  ggplot(aes(name, value)) +
  geom_col() +
  scale_y_continuous(name = "Standard deviation of raw test scores") +
  scale_x_discrete(name = NULL) +
  ggtitle("Standard deviation of test scores over time") +
  theme_few() +
  theme(text = element_text(size = 12))
@
\caption{Standard deviation of test scores across all waves.}
\end{center}
\end{figure}

@@

complete_data_topbottom %>%
  bind_rows(complete_data_topmid) %>%
  bind_rows(complete_data_midbottom) %>%
  filter(country %in% c("United States", "Denmark", "France", "Germany"), type_test == "math") %>%
  select(wave, country, type_test, type, contains("math")) %>%
  mutate(type = factor(type,
                       levels = c("90th/10th SES gap", "90th/50th SES gap", "50th/10th SES gap"),
                       ordered = TRUE)) %>%
  gather(score, value, -(wave:type)) %>%
  separate(score, c("ses", "score"), sep = 2) %>%
  spread(score, value) %>%
  ggplot(aes(as.factor(wave), mean_math, group = ses, colour = ses, shape = ses)) +
  geom_line(alpha = 0.4) +
  geom_point(size = 2, alpha = 0.4) +
  coord_cartesian(ylim = c(-1, 2)) +
  scale_y_continuous(name = "Standardized test scores (mean 0)") +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  scale_colour_discrete(name = NULL, labels = c("Low SES", "High SES")) +
  scale_shape_discrete(name = NULL, labels = c("Low SES", "High SES")) +
  ggtitle("Evolution of the top/bottom groups by different gaps") +
  facet_grid(country ~ type) +
  theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        legend.position = "bottom")
@
\caption{Evolution of the SES gaps for top/bottom groups}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
<<avgperformance_inequality, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
data_to_plot %>%
  ggplot(aes(scale(Mean), difference)) +
  geom_smooth(method = "lm") +
  geom_point() +
  scale_y_continuous(name = "90/10 achievement gap") +
  scale_x_continuous(name = "Average performance standardized") +
  ggtitle("Relationship between average performance and achievement gap") +
  facet_wrap(~ wave, nrow = 2) +
  theme_few()
@
\caption{Relationship between 90/10 achievement gap and average performance in mathematics}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
<<country_achievement_and_inequality, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
data_to_plot %>%
  mutate(Mean = scale(Mean),
         country = factor(country, levels = ordered_cnt, ordered = TRUE)) %>%
  gather(type, value, difference, Mean) %>%
  filter(country %in% countries) %>%
  ggplot(aes(as.character(wave), value, group = type, colour = type, shape = type)) +
  geom_point(size = 2, alpha = 0.4) +
  geom_line(alpha = 0.4) +
  scale_y_continuous(name = "Test scores in SD") +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  scale_colour_discrete(name = NULL, labels = c("Achievement gap", "Avg performance")) +
  scale_shape_discrete(name = NULL, labels = c("Achievement gap", "Avg performance")) +
  ggtitle("Evolution of the top/bottom groups by different gaps") +
  facet_wrap(~ country, ncol = 5) +
  theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        legend.position = "bottom")
@
\caption{Time trends of achievement gap and average achiement between 2000 and 2015}
\end{center}
\end{figure}

### 2017-07-05 paper/paper.tex
Fixed shapes of two graphs

\title{First paper - Draft}
\author{Jorge Cimentada}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\setlength{\parindent}{2em}
\setlength{\parskip}{1em}
\showboxdepth=5
\showboxbreadth=5

@@

\end{center}
\end{figure}

@@

{\centering \includegraphics[width=6.5in,height=5in]{figure/avgperformance_inequality-1}

>>>>

\title{The not so widening achievement gap: an interntational comparison of the evolution of the SES achievement gap from 2000-2015}
\author{Jorge Cimentada}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\setlength{\parindent}{2em}
\setlength{\parskip}{1em}
\showboxdepth=5
\showboxbreadth=5

@@

\caption{Evolution of the SES gaps for top/bottom groups}
\end{center}
\end{figure}

@@

{\centering \includegraphics[width=5.5in,height=5in]{figure/avgperformance_inequality-1}
